---
title: "A tutorial for using the ecodecision package with rasters."
author: "Ian Flint and Roozbeh Valavi"
date: "`r Sys.Date()`"
output: 
  pdf_document
vignette: >
  %\VignetteIndexEntry{tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, tidy.opts=list(width.cutoff=80), fig.align="center", message = FALSE, warning = FALSE, fig.height=5, fig.width=5)
```

## Species distribution modelling


## Envrionmental data

```{r}
library(ecodecision)
library(maptools)
library(raster)
library(sf)
library(spatstat, quietly = TRUE)

remove(list = ls())

set.seed(42)
```

We begin by setting up the data in the raster format.

```{r}
grid_dir <- "data/grids"
vars <- list.files(grid_dir, pattern = ".tif$", full.names = TRUE)

# the border shapefile
ibra <- st_read("data/ibraone.shp", crs = 4283, quiet = TRUE)

# read the raster layers as a raster stack
covariates_raster <- stack(vars)

# set the coordinate system
crs(covariates_raster) <- CRS("+init=epsg:4283")
```

We then convert the data to the format used by `spatstat`.

```{r}
covariates <- lapply(as.list(covariates_raster), 
                     function(element) maptools::as.im.RasterLayer(element))
```

## Species data
The species data comes from Fithian and others (2015).

First, create a `spatstat::owin` that models the observation region.

```{r}
window <- spatstat::as.owin(covariates[[1]])
```

Second, construct the `spatstat::ppp` representing the point configuration.

```{r}
# load Fithian et al (2015) species data
load("data/moddat.RData")

configuration <- spatstat::ppp(eucacine_moddat$long, 
                               eucacine_moddat$lat, 
                               window = window)
```

# Using ecodecision

```{r}
detection_probability <- 0.3
ecodecision::optimal_location(configuration,
                              environmental_covariates = covariates,
                              detection_probability = detection_probability)
```

